---
title: "Support Vector Machine Learning on Later Data"
author: "Lyndsey Umsted"
date: '2022-07-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("caret")
```

```{r}
library(caret)
library(kernlab)
library(ROCR)
```
Importing File into R:
```{r}
aad <- read.csv("data/Amazon_Data_Annual_for_TJ_5_22.csv")
#view(aad)

cutaneous <- aad$Cutaneous.Leishmaniasis
mucosal <- aad$Mucosal.Leishmaniasis
visceral <- aad$Visceral.Leishmaniasis
new_data <- subset(aad, !is.na(cutaneous))
#View(new_data)
#names(new_data)
```


Extracting Later Data and Tidying it:
```{r}
## splitting the data into a before 2014 set and an after 2014 set
library(dplyr)
early_data <- new_data %>%
  filter(Year < 2014)%>%
  select(-c(29:69))
later_data <- new_data %>%
  filter(Year > 2013) %>%
  select(-c(29:69))

#names(early_data)

## removing unnecessary variables

early_data <-gap.inp.early.2$ximp
later_data <- later_data %>%
  select(-c("StableLights"))

# later_data$OptTemp_Obs <- as.numeric(later_data$OptTemp_Obs)
# later_data$Year <- as.numeric(later_data$Year)
# later_data$Population <- as.numeric(later_data$Population)

later_data_small <- later_data %>%
  select(c("Cutaneous.Leishmaniasis", 
           "Dengue_Alb_OptTemp", 
           "Dengue_Aeg_OptTemp", 
           "Chik_Alb_OptTemp", 
           "Chik_Aeg_OptTemp", 
           "Zika_OptTemp", 
           "Malaria_OptTemp", 
           "LST_Day", 
           "LST_Night", 
           "OptTemp_Obs", 
           "NDVI", 
           "EVI", 
           "Precip", 
           "AvgRad", 
           "SWOccurrence")) %>%
  na.omit(later_data_small) 
later_data_t <- later_data_small %>%
  mutate(Cutaneous.Leishmaniasis = (Cutaneous.Leishmaniasis))
```

```{r}
library(dplyr)
cat_df <- subset(early_data, early_data$Cutaneous.Leishmaniasis > 0)
cat_df$Cutaneous.Leishmaniasis <- cut(early_data$Cutaneous.Leishmaniasis[early_data$Cutaneous.Leishmaniasis > 0], breaks = c(0, 0.09154 , 0.86616, 100), labels = c("low", "moderate", "high")) # 25%, 75%, 100%

cat_df <- cat_df %>%
  select(c("Cutaneous.Leishmaniasis", 
           "Dengue_Alb_OptTemp", 
           "Dengue_Aeg_OptTemp", 
           "Chik_Alb_OptTemp", 
           "Chik_Aeg_OptTemp", 
           "Zika_OptTemp", 
           "Malaria_OptTemp", 
           "LST_Day", 
           "LST_Night", 
           "OptTemp_Obs", 
           "NDVI", 
           "EVI", 
           "Precip", 
           "StableLights", 
           "SWOccurrence")) %>%
  na.omit(cat_df) 

# cat_df$label <- NA
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "low"] <- 0
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "moderate"] <- 1
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "high"] <- 2

# cat_df$Year <- as.numeric(cat_df$Year)
# cat_df$Population <- as.numeric(cat_df$Population)
# cat_df$OptTemp_Obs <- as.numeric(cat_df$OptTemp_Obs)
```

```{r}
skimr::skim(cat_df)
round(prop.table(table(cat_df$Cutaneous.Leishmaniasis)), 2)
```

```{r}
str(cat_df)
```

Partitioning Data into Training and Test Sets
```{r}
set.seed(1)
library(rsample)
data_split <- initial_split(cat_df, strata = "Cutaneous.Leishmaniasis", prop = 0.8)

training <- training(data_split) %>%
  group_by(Cutaneous.Leishmaniasis) %>%
  sample_n(size = 6377)
testing <- testing(data_split) 

x_train <- as.matrix(training[,-1])
x_test <- as.matrix(testing[,-1])

y_train <- as.vector(as.numeric(training$Cutaneous.Leishmaniasis) - 1)
y_test <- as.vector(as.numeric(testing$Cutaneous.Leishmaniasis) - 1)
 
# training <- subset(training, select = -Cutaneous.Leishmaniasis)
# testing <- subset(testing, select = -Cutaneous.Leishmaniasis)

```

```{r}
summary(cat_df)
```

Factorize our target variable
```{r}
training[["Cutaneous.Leishmaniasis"]] = factor(training[["Cutaneous.Leishmaniasis"]])
```

traincontrol() method:
```{r, eval = FALSE}
# 
set.seed(2022)
trctrl <- trainControl(method = "cv", number = 2, classProbs = TRUE)
# 
trctrl_1 <- trainControl(method = 'cv', number = 1)
# 
# # grid search to fine tune
grid <- expand.grid(C = c(0,0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2,5))
# 
# # train SVM
# svm_Radial <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmRadial",
#                     trControl=trctrl,
#                     preProcess = c("center", "scale"),
#                     tuneLength = 10)
# 
svm_Linear <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmLinear",
                    trControl=trctrl,
                    preProcess = c("center", "scale"),
                    tuneLength = 10)
# svm_Poly <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmPoly",
#                     trControl=trctrl_1,
#                     preProcess = c("center", "scale"),
#                     tuneLength = 10)
system("say Just finished!")
# alarm()
# save(svm_Linear, file = 'models/svm_Linear')
# save(svm_Radial, file = 'models/svm_Radial')
```

```{r}
# library(e1071)
# svmfit_poly = svm(Cutaneous.Leishmaniasis ~ ., data = training, kernel = 'radial', gamma = 1, cost = 1)
# save(svmfit_poly, file = 'models/svmfit_poly')
```

```{r}
# set.seed(1)
# tune.out = tune(svm, Cutaneous.Leishmaniasis ~ ., 
#                 data = training, kernel = 'radial',
#                 ranges = list(cost = c(0.1, 1, 10, 100, 1000),
#                               gamma = c(0.5, 1, 2, 3, 4)))
# save(tune.out, file = 'models/tune.out')
```

```{r}
load(file = 'models/svm_Linear')
load(file = 'models/svm_Radial')
load(file = 'models/tune.out')



confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
      pred = predict(tune.out$best.model, newdata = testing %>%
                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
```


```{r}
table(true = testing$Cutaneous.Leishmaniasis,
      pred = predict(tune.out$best.model, newdata = testing %>%
                       dplyr::select(-c('Cutaneous.Leishmaniasis'))))
plot(tune.out)

```

```{r}
# Do not run anymore!!! Long runtime
# tune.out.2 = tune(svm, Cutaneous.Leishmaniasis ~ ., 
#                 data = training, kernel = 'radial',
#                 ranges = list(cost = c(1, 2, 4, 7, 10, 11, 12, 14, 16),
#                               sigma = c(0.01, 0.02, 0.04, 0.06, 
#                                         0.08, 0.1, 0.2, 0.3),
#                               gamma = c(0.01, 0.25, 0.5, 0.75)))

# save(tune.out.2, file = 'models/tune.out.2')

```




```{r}
test_pred_linear <- predict(svm_Linear, newdata = testing)
test_pred_radial <- predict(svm_Radial, newdata = testing)
```

Predicting Accuracy with Confusion Matrix
```{r}
confusionMatrix(table(test_pred_linear, testing$Cutaneous.Leishmaniasis))
confusionMatrix(table(test_pred_radial, testing$Cutaneous.Leishmaniasis))
```

We are going to put some values of C using expand.grid() into “grid” dataframe. Next step is to use this dataframe for testing our classifier at specific C values. It needs to be put in train() method with tuneGrid parameter.
```{r}
grid <- expand.grid(C = c(10, 11, 12, 13, 14, 15, 16, 17),
                    sigma = c(0.01, 0.02, 0.03, 0.04, 0.05, 0.06, 0.07, 0.08, 0.09, 0.1))
grid_Linear <- expand.grid(C = c(10, 11, 12, 13, 14, 15, 16, 17))
svm_Linear_Grid <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmLinear",
                         trControl=trctrl,
                         preProcess = c("center", "scale"),
                         tuneGrid = grid_Linear,
                         tuneLength = 10)
svm_Radial_Grid <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmRadial",
                         trControl=trctrl,
                         preProcess = c("center", "scale"),
                         tuneGrid = grid,
                         tuneLength = 10)


save(svm_Linear_Grid, file = 'models/svm_Linear_Grid_imp_2')
save(svm_Radial_Grid, file = 'models/svm_Radial_Grid_imp_2')
# 
# plot(svm_Linear_Grid)
# plot(svm_Radial_Grid)
```

Let’s try to make predictions using this model for our test set.
```{r}
test_pred_grid_linear <- predict(svm_Linear_Grid, newdata = testing)
test_pred_grid_radial <- predict(svm_Radial_Grid, newdata = testing)
```

Confusion Matrix:
```{r}
confusionMatrix(table(test_pred_grid_linear, testing$Cutaneous.Leishmaniasis))
confusionMatrix(table(test_pred_grid_radial, testing$Cutaneous.Leishmaniasis))
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
      pred = predict(tune.out.2$best.model, newdata = testing %>%
                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
```




```{r}
# load(file = 'models/svm_Linear')
# load(file = 'models/svm_Radial')
# load(file = 'models/svmfit_poly')
# load(file = 'models/tune.out')
# load(file = 'models/tune.out.2')
# load(file - 'models/svm_Linear_Grid')
# load(file = 'models/svm_Radial_Grid')`


summary(svm_Linear)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(svm_Linear, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
summary(svm_Radial)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(svm_Radial, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
summary(tune.out)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(tune.out$best.model, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
summary(tune.out.2)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(tune.out.2$best.model, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
summary(svm_Linear_Grid)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(svm_Linear_Grid, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))
summary(svm_Radial_Grid)
confusionMatrix(table(true = testing$Cutaneous.Leishmaniasis,
                      pred = predict(svm_Radial_Grid, newdata = testing %>%
                                       dplyr::select(-c('Cutaneous.Leishmaniasis')))))


```






```{python}
import pandas as pd
```

```{python}
toy = pd.DataFrame(a = {})
```

