---
title: "Stacked Model for Cutaneous Leishmaniasis"
author: "Lyndsey Umsted"
date: "2022-07-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Neccessary Libraries
```{r}
library(tidyverse)
library(tidymodels)
library(stacks)
library(pROC)
```

Loading the Annual Amazon Data Set and Tidying it by removing any NA values for Cutaneuous Leishmaniasis along with any case reportings of 0.
```{r}
aad <- read.csv("data/aad.csv")
aad <- subset(aad, !is.na(aad$Cutaneous.Leishmaniasis))
aad <-subset(aad, aad$Cutaneous.Leishmaniasis > 0)
```

Separating the data into two separate data sets, one for data before 2014 (early_data) and one for data after 2013 (later_data), and then tidying it while removing unnecceassary columns. 
```{r}
library(dplyr)
early_data <- aad %>%
  filter(Year < 2014)
later_data <- aad %>%
  filter(Year > 2013)

early_data <- early_data %>%
  select(-c("AvgRad"))
later_data <- later_data %>%
  select(-c("StableLights"))
```

Reducing columns to desired Predicitive Variables: Population, LAST_Day, Precip, AvgRad, SWOccurrence, NDVI, EVI, pland_forest, te_forest, enn_mn_forest.
```{r}
later_data_small <- later_data %>%
  select(c("Population", "Cutaneous.Leishmaniasis", "LST_Day", "Precip", "AvgRad", "SWOccurrence", "NDVI", "EVI", "pland_forest", "te_forest", "enn_mn_forest"))
later_data_small <- subset(later_data_small, !is.na(later_data_small$LST_Day))
later_data_small <- subset(later_data_small, !is.na(later_data_small$SWOccurrence))
```

Converting all NA values in pland_forest, te_forest, and enn_mn_forest.
```{r}
later_data_small$pland_forest <- ifelse(is.na(later_data_small$pland_forest), 
                                        0, later_data_small$pland_forest)
later_data_small$te_forest <- ifelse(is.na(later_data_small$te_forest), 
                                     0, later_data_small$te_forest)
later_data_small$enn_mn_forest <- ifelse(is.na(later_data_small$enn_mn_forest), 
                                         0, later_data_small$enn_mn_forest)
```

Creating new dataset cat_df where values of Cutaneous.Leishmaniasis are split at the 30th percentile threshold and categorized into low and high risk labels.
```{r}
library(dplyr)
summary(later_data_small$Cutaneous.Leishmaniasis)
cat_df <- later_data_small
cat_df$Cutaneous.Leishmaniasis <- cut(later_data_small$Cutaneous.Leishmaniasis, 
                                      breaks = c(0, 0.11703435, 10^3), 
                                      labels = c("low", "high")) # median


skimr::skim(cat_df)
round(prop.table(table(cat_df$Cutaneous.Leishmaniasis)), 2)
```

Creating Testing and Training sets for our base models.
```{r}
set.seed(123) # for reproducibility

split <- initial_split(cat_df)

data_train <- training(split)
data_test <- testing(split)

# use a 5-fold cross-validation
folds <- rsample::vfold_cv(data_train, v = 5)
```

Creating basic recipe defining both predictor and target variables, along with definig a minimal workflow and metric for all the base models.
```{r}
# set up a basic recipe
data_rec <-
  recipe(Cutaneous.Leishmaniasis ~ Population + LST_Day + NDVI + 
           EVI + Precip + AvgRad + SWOccurrence + pland_forest + 
           te_forest + enn_mn_forest, data = data_train) %>%
  step_dummy(all_nominal() - all_outcomes()) %>% 
  step_zv(all_predictors())

# define a minimal workflow
data_wflow <-
  workflow() %>% 
  add_recipe(data_rec)

# add metric rmse (same as malaria)
metric <- metric_set(recall, accuracy) #.784 for roc_auc
```

Setting up control grids and resamples along with a logistic regression model for
```{r}
library(stacks)

# save assessment set predictions and workflow used to fit the resamples
ctrl_grid <- control_stack_grid()
ctrl_res <- control_stack_resamples()

# models: SVM, XGBoost, RF
## models to try: logistic regression 

# toy model
log_reg_spec <-
  logistic_reg() %>%
  set_engine('glm')

log_reg_wflow <- 
  data_wflow %>%
  add_model(log_reg_spec)

set.seed(123)
log_reg_res <-
  fit_resamples(
    log_reg_wflow,
    resamples = folds,
    metrics = metric,
    control = ctrl_res
  )
```

Setting up Support Vector Machine base model:
```{r}
# define svm model using parsnip
svm_spec <-
  svm_rbf(
    cost = tune(),
    rbf_sigma = tune()
  ) %>%
  set_engine('kernlab') %>%
  set_mode('classification')

# add it to a workflow
svm_wflow <-
  data_wflow %>%
  add_model(svm_spec)

# tune cost and rbf_sigma and fit to the 10-fold cv
set.seed(123)
svm_res <-
  tune_grid(
    svm_wflow,
    resamples = folds,
    grid = 5,
    control = ctrl_grid
  )

save(svm_res, file = 'stacking/svm_res_later_classif_2')
#load(file = 'models/stacking/svm_res_later_classif_2')

```

Setting up XGBoost base model:
```{r}
# define xgboost model using parsnip

set.seed(123)
xgb_spec <-
  boost_tree(
    mtry = tune(),
    trees = tune(),
    min_n = tune(),
    tree_depth = tune(),
    learn_rate = tune(),
    loss_reduction = tune()
  ) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

# add it to a workflow
xgb_wflow <-
  data_wflow %>%
  add_model(xgb_spec)

# tune mtry, trees, min_n, tree_depth, etc.
xgb_res <-
  tune_grid(
    xgb_wflow,
    resamples = folds,
    grid = 5,
    control = ctrl_grid
  )

save(xgb_res, file = 'stacking/xgb_res_later_classif_2')

#load(file = 'models/stacking/xgb_res_later_classif_2')
```

Setting up Random Forest base model:
```{r}
# define rf model using parsnip

set.seed(123)
rf_spec <-
  rand_forest(
    mtry = tune(),
    trees = tune(),
    min_n = tune()
  ) %>%
  set_engine('ranger') %>%
  set_mode('classification')

# add it to a workflow
rf_wflow <-
  data_wflow %>%
  add_model(rf_spec)

# tune mtry, trees, min_n
rf_res <-
  tune_grid(
    rf_wflow,
    resamples = folds,
    grid = 5,
    control = ctrl_grid
  )

save(rf_res, file = 'stacking/rf_res_later_classif_2')

#load(file = 'models/stacking/rf_res_later_classif_2')
```

Adding candidate models to stacked model
```{r}
data_st <- 
  stacks() %>% 
  add_candidates(svm_res) %>% 
  add_candidates(xgb_res) %>% 
  add_candidates(rf_res)

data_st

as_tibble(data_st)
```

Blending base model predictions and fitting them.
```{r}
model_st <-
  data_st %>% 
  blend_predictions()

model_st <- 
  model_st %>% 
  fit_members()
```

Using stacked model to make predictions on Cutaneous Leishmaniasis and Preparing Confusion Matrix
```{r}
data_test <-
  data_test %>%
  bind_cols(predict(model_st, .))

# confusion matrix for stacks
caret::confusionMatrix(data_test$Cutaneous.Leishmaniasis, 
                       data_test$.pred_class,
                       positive = 'high')
```

Plot of positive predictive and negative predictive values for Cutaneous Leishmaniasis at 30th percentile threshold split.
```{r}
plot(data_test$Cutaneous.Leishmaniasis, data_test$.pred_class)
```
Looking at the Separate Accurracies of each base model in the stacked model.
```{r}
# confusion matrix for base models

member_preds <- 
  data_test %>% 
  select(Cutaneous.Leishmaniasis) %>% 
  bind_cols(
    predict(
      model_st,
      data_test,
      members = TRUE
    )
  )


colnames(member_preds) %>% 
  map_dfr(
    .f = accuracy,
    truth = Cutaneous.Leishmaniasis,
    data = member_preds
  ) %>% 
  mutate(member = colnames(member_preds))
```





