---
title: "Data Exploration"
author: "Lyndsey Umsted"
date: '2022-07-11'
output: html_document
---

```{r}
aad <- read.csv("data/aad.csv")
# removed_na <- subset(aad, !is.na(aad$Cutaneous.Leishmaniasis))
# missing_to_zero <- aad
# missing_to_zero$Cutaneous.Leishmaniasis[is.na(missing_to_zero$Cutaneous.Leishmaniasis)] <- 0
# View(aad)
library(dplyr)

early_data <- aad %>%
  filter(Year < 2014)%>%
  dplyr::select(c(1:28,30,36,40,45))
later_data <- aad %>%
  filter(Year > 2013) %>%
  dplyr::select(c(1:28,30,36,40,45))

#names(early_data)

## removing unnecessary variables

early_data <- early_data %>%
  dplyr::select(-c("AvgRad"))
later_data <- later_data %>%
  dplyr::select(-c("StableLights")) 

# early_data <- subset(early_data, !is.na(c(early_data$pland_forest, early_data$area_mn_forest, early_data$te_forest,early_data$enn_mn_forest, early_data$SWOccurrence, early_data$StableLights)))
# later_data <- subset(later_data, !is.na(c(later_data$pland_forest, later_data$area_mn_forest, later_data$te_forest,later_data$enn_mn_forest, later_data$SWOccurrence, later_data$AvgRad)))
```

```{r}
boxplot(removed_na$enn_mn_riverlakeocean)
```


```{r}
hist(removed_na$area_mn_forest)
```

```{r}
library(dplyr)

early_data <- early_data %>%
  dplyr::select(-c("Country", "Name"))
later_data <- later_data %>%
  dplyr::select(-c("Country", "Name")) 

without_na_pop <- subset(early_data, !is.na(early_data$Population))
mean_population <- early_data %>%
  group_by(Year) %>%
  summarise_at(vars(Population), list(name = mean))
mean_population

```


## EARLY DATA

Population:
```{r}
without_na_pop <- subset(early_data, !is.na(early_data$Population))
mean_population <- without_na_pop %>%
  group_by(Year) %>%
  summarise_at(vars(Population), list(name = mean))
mean_population
```

Chikungunya: (ALL NAs)
```{r}
without_na_chik <- subset(early_data, !is.na(early_data$Chikungunya))
mean_chik <- without_na_chik %>%
  group_by(Year) %>%
  summarise_at(vars(Chikungunya), list(name = mean))
mean_chik
```

Cutaneous.Leishmaniasis
```{r}
without_na_Cutaneous <- subset(early_data, !is.na(early_data$Cutaneous))
mean_Cutaneous <- without_na_Cutaneous %>%
  group_by(Year) %>%
  summarise_at(vars(Cutaneous.Leishmaniasis), list(name = mean))
mean_Cutaneous
```

Dengue:
```{r}
without_na_Dengue <- subset(early_data, !is.na(early_data$Dengue))
mean_Dengue <- without_na_Dengue %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue), list(name = mean))
mean_Dengue
```

Malaria:
```{r}
without_na_Malaria <- subset(early_data, !is.na(early_data$Malaria))
mean_Malaria <- without_na_Malaria %>%
  group_by(Year) %>%
  summarise_at(vars(Malaria), list(name = mean))
mean_Malaria
```

Mucosal.Leishmaniasis

```{r}
without_na_Mucosal.Leishmaniasis <- subset(early_data, !is.na(early_data$Mucosal.Leishmaniasis))
mean_Mucosal.Leishmaniasis <- without_na_Mucosal.Leishmaniasis %>%
  group_by(Year) %>%
  summarise_at(vars(Mucosal.Leishmaniasis), list(name = mean))
mean_Mucosal.Leishmaniasis
```

Visceral.Leishmaniasis:
```{r}
without_na_Visceral.Leishmaniasis <- subset(early_data, !is.na(early_data$Visceral.Leishmaniasis))
mean_Visceral.Leishmaniasis <- without_na_Visceral.Leishmaniasis %>%
  group_by(Year) %>%
  summarise_at(vars(Visceral.Leishmaniasis), list(name = mean))
mean_Visceral.Leishmaniasis
```


Yellow.Fever
```{r}
without_na_Yellow.Fever <- subset(early_data, !is.na(early_data$Yellow.Fever))
mean_Yellow.Fever <- without_na_Yellow.Fever %>%
  group_by(Year) %>%
  summarise_at(vars(Yellow.Fever), list(name = mean))
mean_Yellow.Fever
```

Zika:
```{r}
without_na_Zika <- subset(early_data, !is.na(early_data$Zika))
mean_Zika <- without_na_Zika %>%
  group_by(Year) %>%
  summarise_at(vars(Zika), list(name = mean))
mean_Zika
```

LST_Day:
```{r}
without_na_LST_Day <- subset(early_data, !is.na(early_data$LST_Day))
mean_LST_Day <- without_na_LST_Day %>%
  group_by(Year) %>%
  summarise_at(vars(LST_Day), list(name = mean))
mean_LST_Day
```

LST_Night:
```{r}
without_na_LST_Night <- subset(early_data, !is.na(early_data$LST_Night))
mean_LST_Night <- without_na_LST_Night %>%
  group_by(Year) %>%
  summarise_at(vars(LST_Night), list(name = mean))
mean_LST_Night
```

OptTemp_Obs:
```{r}
without_na_OptTemp_Obs <- subset(early_data, !is.na(early_data$OptTemp_Obs))
mean_OptTemp_Obs <- without_na_OptTemp_Obs %>%
  group_by(Year) %>%
  summarise_at(vars(OptTemp_Obs), list(name = mean))
mean_OptTemp_Obs
```

Dengue_Alb_OptTemp:
```{r}
without_na_Dengue_Alb_OptTemp <- subset(early_data, !is.na(early_data$Dengue_Alb_OptTemp))
mean_Dengue_Alb_OptTemp <- without_na_Dengue_Alb_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue_Alb_OptTemp), list(name = mean))
mean_Dengue_Alb_OptTemp
```

Dengue_Aeg_OptTemp:
```{r}
without_na_Dengue_Aeg_OptTemp <- subset(early_data, !is.na(early_data$Dengue_Aeg_OptTemp))
mean_Dengue_Aeg_OptTemp <- without_na_Dengue_Aeg_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue_Aeg_OptTemp), list(name = mean))
mean_Dengue_Aeg_OptTemp
```

Chik_Alb_OptTemp:
```{r}
without_na_Chik_Alb_OptTemp <- subset(early_data, !is.na(early_data$Chik_Alb_OptTemp))
mean_Chik_Alb_OptTemp <- without_na_Chik_Alb_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Chik_Alb_OptTemp), list(name = mean))
mean_Chik_Alb_OptTemp
```

Chik_Aeg_OptTemp:
```{r}
without_na_Chik_Aeg_OptTemp <- subset(early_data, !is.na(early_data$Chik_Aeg_OptTemp))
mean_Chik_Aeg_OptTemp <- without_na_Chik_Aeg_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Chik_Aeg_OptTemp), list(name = mean))
mean_Chik_Aeg_OptTemp
```

Zika_OptTemp:
```{r}
without_na_Zika_OptTemp <- subset(early_data, !is.na(early_data$Zika_OptTemp))
mean_Zika_OptTemp <- without_na_Zika_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Zika_OptTemp), list(name = mean))
mean_Zika_OptTemp
```

Malaria_OptTemp:
```{r}
without_na_Malaria_OptTemp <- subset(early_data, !is.na(early_data$Malaria_OptTemp))
mean_Malaria_OptTemp <- without_na_Malaria_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Malaria_OptTemp), list(name = mean))
mean_Malaria_OptTemp
```

NDVI:
```{r}
without_na_NDVI <- subset(early_data, !is.na(early_data$NDVI))
mean_NDVI <- without_na_NDVI %>%
  group_by(Year) %>%
  summarise_at(vars(NDVI), list(name = mean))
mean_NDVI
```

EVI:
```{r}
without_na_EVI <- subset(early_data, !is.na(early_data$EVI))
mean_EVI <- without_na_EVI %>%
  group_by(Year) %>%
  summarise_at(vars(EVI), list(name = mean))
mean_EVI
```

Precip:
```{r}
without_na_Precip <- subset(early_data, !is.na(early_data$Precip))
mean_Precip <- without_na_Precip %>%
  group_by(Year) %>%
  summarise_at(vars(Precip), list(name = mean))
mean_Precip
```

StableLights:
```{r}
without_na_StableLights <- subset(early_data, !is.na(early_data$StableLights))
mean_StableLights <- without_na_StableLights %>%
  group_by(Year) %>%
  summarise_at(vars(StableLights), list(name = mean))
mean_StableLights
```

SWOccurrence
```{r}
without_na_SWOccurrence <- subset(early_data, !is.na(early_data$SWOccurrence))
mean_SWOccurrence <- without_na_SWOccurrence %>%
  group_by(Year) %>%
  summarise_at(vars(SWOccurrence), list(name = mean))
mean_SWOccurrence
```

pland_forest:
```{r}
without_na_pland_forest <- subset(early_data, !is.na(early_data$pland_forest))
mean_pland_forest <- without_na_pland_forest %>%
  group_by(Year) %>%
  summarise_at(vars(pland_forest), list(name = mean))
mean_pland_forest
```

area_mn_forest:
```{r}
without_na_area_mn_forest <- subset(early_data, !is.na(early_data$area_mn_forest))
mean_area_mn_forest <- without_na_area_mn_forest %>%
  group_by(Year) %>%
  summarise_at(vars(area_mn_forest), list(name = mean))
mean_area_mn_forest
```

te_forest:
```{r}
without_na_te_forest <- subset(early_data, !is.na(early_data$te_forest))
mean_te_forest <- without_na_te_forest %>%
  group_by(Year) %>%
  summarise_at(vars(te_forest), list(name = mean))
mean_te_forest
```

enn_mn_forest:
```{r}
without_na_enn_mn_forest <- subset(early_data, !is.na(early_data$enn_mn_forest))
mean_enn_mn_forest <- without_na_enn_mn_forest %>%
  group_by(Year) %>%
  summarise_at(vars(enn_mn_forest), list(name = mean))
mean_enn_mn_forest
```



## LATER DATA


Population:
```{r}
without_na_pop <- subset(later_data, !is.na(later_data$Population))
mean_population2 <- without_na_pop %>%
  group_by(Year) %>%
  summarise_at(vars(Population), list(name = mean))
mean_population2
```

Chikungunya: (ALL NAs)
```{r}
without_na_chik <- subset(later_data, !is.na(later_data$Chikungunya))
mean_chik2 <- without_na_chik %>%
  group_by(Year) %>%
  summarise_at(vars(Chikungunya), list(name = mean))
mean_chik2
```

Cutaneous.Leishmaniasis
```{r}
without_na_Cutaneous <- subset(later_data, !is.na(later_data$Cutaneous))
mean_Cutaneous2 <- without_na_Cutaneous %>%
  group_by(Year) %>%
  summarise_at(vars(Cutaneous.Leishmaniasis), list(name = mean))
mean_Cutaneous2
```

Dengue:
```{r}
without_na_Dengue <- subset(later_data, !is.na(later_data$Dengue))
mean_Dengue2 <- without_na_Dengue %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue), list(name = mean))
mean_Dengue2
```

Malaria:
```{r}
without_na_Malaria <- subset(later_data, !is.na(later_data$Malaria))
mean_Malaria2 <- without_na_Malaria %>%
  group_by(Year) %>%
  summarise_at(vars(Malaria), list(name = mean))
mean_Malaria2
```

Mucosal.Leishmaniasis

```{r}
without_na_Mucosal.Leishmaniasis <- subset(later_data, !is.na(later_data$Mucosal.Leishmaniasis))
mean_Mucosal.Leishmaniasis2 <- without_na_Mucosal.Leishmaniasis %>%
  group_by(Year) %>%
  summarise_at(vars(Mucosal.Leishmaniasis), list(name = mean))
mean_Mucosal.Leishmaniasis2
```

Visceral.Leishmaniasis:
```{r}
without_na_Visceral.Leishmaniasis <- subset(later_data, !is.na(later_data$Visceral.Leishmaniasis))
mean_Visceral.Leishmaniasis2 <- without_na_Visceral.Leishmaniasis %>%
  group_by(Year) %>%
  summarise_at(vars(Visceral.Leishmaniasis), list(name = mean))
mean_Visceral.Leishmaniasis2
```


Yellow.Fever
```{r}
without_na_Yellow.Fever <- subset(later_data, !is.na(later_data$Yellow.Fever))
mean_Yellow.Fever2 <- without_na_Yellow.Fever %>%
  group_by(Year) %>%
  summarise_at(vars(Yellow.Fever), list(name = mean))
mean_Yellow.Fever2
```

Zika:
```{r}
without_na_Zika <- subset(later_data, !is.na(later_data$Zika))
mean_Zika2 <- without_na_Zika %>%
  group_by(Year) %>%
  summarise_at(vars(Zika), list(name = mean))
mean_Zika2
```

LST_Day:
```{r}
without_na_LST_Day <- subset(later_data, !is.na(later_data$LST_Day))
mean_LST_Day2 <- without_na_LST_Day %>%
  group_by(Year) %>%
  summarise_at(vars(LST_Day), list(name = mean))
mean_LST_Day2
```

LST_Night:
```{r}
without_na_LST_Night <- subset(later_data, !is.na(later_data$LST_Night))
mean_LST_Night2 <- without_na_LST_Night %>%
  group_by(Year) %>%
  summarise_at(vars(LST_Night), list(name = mean))
mean_LST_Night2
```

OptTemp_Obs:
```{r}
without_na_OptTemp_Obs <- subset(later_data, !is.na(later_data$OptTemp_Obs))
mean_OptTemp_Obs2 <- without_na_OptTemp_Obs %>%
  group_by(Year) %>%
  summarise_at(vars(OptTemp_Obs), list(name = mean))
mean_OptTemp_Obs2
```

Dengue_Alb_OptTemp:
```{r}
without_na_Dengue_Alb_OptTemp <- subset(later_data, !is.na(later_data$Dengue_Alb_OptTemp))
mean_Dengue_Alb_OptTemp2 <- without_na_Dengue_Alb_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue_Alb_OptTemp), list(name = mean))
mean_Dengue_Alb_OptTemp2
```

Dengue_Aeg_OptTemp:
```{r}
without_na_Dengue_Aeg_OptTemp <- subset(later_data, !is.na(later_data$Dengue_Aeg_OptTemp))
mean_Dengue_Aeg_OptTemp2 <- without_na_Dengue_Aeg_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Dengue_Aeg_OptTemp), list(name = mean))
mean_Dengue_Aeg_OptTemp2
```

Chik_Alb_OptTemp:
```{r}
without_na_Chik_Alb_OptTemp <- subset(later_data, !is.na(later_data$Chik_Alb_OptTemp))
mean_Chik_Alb_OptTemp2 <- without_na_Chik_Alb_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Chik_Alb_OptTemp), list(name = mean))
mean_Chik_Alb_OptTemp2
```

Chik_Aeg_OptTemp:
```{r}
without_na_Chik_Aeg_OptTemp <- subset(later_data, !is.na(later_data$Chik_Aeg_OptTemp))
mean_Chik_Aeg_OptTemp2 <- without_na_Chik_Aeg_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Chik_Aeg_OptTemp), list(name = mean))
mean_Chik_Aeg_OptTemp2
```

Zika_OptTemp:
```{r}
without_na_Zika_OptTemp <- subset(later_data, !is.na(later_data$Zika_OptTemp))
mean_Zika_OptTemp2 <- without_na_Zika_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Zika_OptTemp), list(name = mean))
mean_Zika_OptTemp2
```

Malaria_OptTemp:
```{r}
without_na_Malaria_OptTemp <- subset(later_data, !is.na(later_data$Malaria_OptTemp))
mean_Malaria_OptTemp2 <- without_na_Malaria_OptTemp %>%
  group_by(Year) %>%
  summarise_at(vars(Malaria_OptTemp), list(name = mean))
mean_Malaria_OptTemp2
```

NDVI:
```{r}
without_na_NDVI <- subset(later_data, !is.na(later_data$NDVI))
mean_NDVI2 <- without_na_NDVI %>%
  group_by(Year) %>%
  summarise_at(vars(NDVI), list(name = mean))
mean_NDVI2
```

EVI:
```{r}
without_na_EVI <- subset(later_data, !is.na(later_data$EVI))
mean_EVI2 <- without_na_EVI %>%
  group_by(Year) %>%
  summarise_at(vars(EVI), list(name = mean))
mean_EVI2
```

Precip:
```{r}
without_na_Precip <- subset(later_data, !is.na(later_data$Precip))
mean_Precip2 <- without_na_Precip %>%
  group_by(Year) %>%
  summarise_at(vars(Precip), list(name = mean))
mean_Precip2
```

AvgRad:
```{r}
without_na_AvgRad <- subset(later_data, !is.na(later_data$AvgRad))
mean_AvgRad <- without_na_AvgRad %>%
  group_by(Year) %>%
  summarise_at(vars(AvgRad), list(name = mean))
mean_AvgRad
```


SWOccurrence
```{r}
without_na_SWOccurrence <- subset(early_data, !is.na(early_data$SWOccurrence))
mean_SWOccurrence2 <- without_na_SWOccurrence %>%
  group_by(Year) %>%
  summarise_at(vars(SWOccurrence), list(name = mean))
mean_SWOccurrence2
```

pland_forest:
```{r}
without_na_pland_forest <- subset(later_data, !is.na(later_data$pland_forest))
mean_pland_forest2 <- without_na_pland_forest %>%
  group_by(Year) %>%
  summarise_at(vars(pland_forest), list(name = mean))
mean_pland_forest2
```

area_mn_forest:
```{r}
without_na_area_mn_forest <- subset(later_data, !is.na(later_data$area_mn_forest))
mean_area_mn_forest2 <- without_na_area_mn_forest %>%
  group_by(Year) %>%
  summarise_at(vars(area_mn_forest), list(name = mean))
mean_area_mn_forest2
```

te_forest:
```{r}
without_na_te_forest <- subset(later_data, !is.na(later_data$te_forest))
mean_te_forest2 <- without_na_te_forest %>%
  group_by(Year) %>%
  summarise_at(vars(te_forest), list(name = mean))
mean_te_forest2
```

enn_mn_forest:
```{r}
without_na_enn_mn_forest <- subset(later_data, !is.na(later_data$enn_mn_forest))
mean_enn_mn_forest2 <- without_na_enn_mn_forest %>%
  group_by(Year) %>%
  summarise_at(vars(enn_mn_forest), list(name = mean))
mean_enn_mn_forest2
```

```{r}
library(tidyverse)

df_list <- list(mean_population, mean_chik, mean_Cutaneous, mean_Dengue, mean_Malaria, mean_Mucosal.Leishmaniasis, mean_Visceral.Leishmaniasis, mean_Yellow.Fever, mean_Zika, mean_LST_Day, mean_LST_Night, mean_OptTemp_Obs, mean_Dengue_Alb_OptTemp, mean_Dengue_Aeg_OptTemp, mean_Chik_Alb_OptTemp, mean_Chik_Aeg_OptTemp, mean_Zika_OptTemp, mean_Malaria_OptTemp, mean_NDVI, mean_EVI, mean_Precip, mean_StableLights, mean_SWOccurrence, mean_pland_forest, mean_area_mn_forest, mean_te_forest, mean_enn_mn_forest)

df_list %>% Reduce(full_join, by="Year")
```

```{r}
mean_population$Population <- mean_population$name
mean_chik$Chikungunya <- mean_chik$name
mean_Cutaneous$Cutaneous.Lesihmaniasis <- mean_Cutaneous$name
mean_Dengue$Dengue <- mean_Dengue$name
mean_Malaria$Malaria <- mean_Malaria$name
mean_Mucosal.Leishmaniasis$Mucosal.Leishmaniasis <- mean_Mucosal.Leishmaniasis$name
mean_Visceral.Leishmaniasis$Visceral.Leishmaniasis <- mean_Visceral.Leishmaniasis$name
mean_Yellow.Fever$Yellow.Fever <- mean_Yellow.Fever$name
mean_Zika$Zika <- mean_Zika$name
mean_LST_Day$LST_Day <- mean_LST_Day$name
mean_LST_Night$LST_Night <- mean_LST_Night$name
mean_OptTemp_Obs$OptTemp_Obs <- mean_OptTemp_Obs$name
mean_Dengue_Alb_OptTemp$Dengue_Alb_OptTemp  <- mean_Dengue_Alb_OptTemp$name
mean_Dengue_Aeg_OptTemp$Dengue_Aeg_OptTemp <- mean_Dengue_Aeg_OptTemp$name
mean_Chik_Alb_OptTemp$Chik_Alb_OptTemp <- mean_Chik_Alb_OptTemp$name
mean_Chik_Aeg_OptTemp$Chik_Aeg_OptTemp <- mean_Chik_Aeg_OptTemp$name
mean_Zika_OptTemp$Zika_OptTemp <- mean_Zika_OptTemp$name
mean_Malaria_OptTemp$Malaria_OptTemp <- mean_Malaria_OptTemp$name
mean_NDVI$NDVI <- mean_NDVI$name
mean_EVI$EVI <- mean_EVI$name
mean_Precip$Precip <- mean_Precip$name
mean_StableLights$StableLights <- mean_StableLights$name
mean_SWOccurrence$SWOccurrence <- mean_SWOccurrence$name
mean_pland_forest$pland_forest <- mean_pland_forest$name
mean_area_mn_forest$area_mn_forest <- mean_area_mn_forest$name
mean_te_forest$te_forest <- mean_te_forest$name
mean_enn_mn_forest$enn_mn_forest <- mean_enn_mn_forest$name
```


```{r}
library(plyr)
new_df <- cbind(mean_population, mean_chik)
new_df
```