---
title: "Support Vector Machine Learning on Later Data"
author: "Lyndsey Umsted"
date: '2022-07-01'
output: html_document
---

```{r}
HFI <- read.csv("data/HFI.csv")
```



```{r}
library(dplyr)
summary(HFI$c_leish_incidence[HFI$c_leish_incidence > 0])
cat_df <- subset(HFI, HFI$c_leish_incidence > 0)
cat_df$c_leish_incidence <- cut(HFI$c_leish_incidence[HFI$c_leish_incidence > 0], breaks = c(0, 0.08837867, 0.34495257, 10^3), labels = c("low", "moderate", "high")) # 33%, 67%, 100%


# cat_df$label <- NA
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "low"] <- 0
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "moderate"] <- 1
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "high"] <- 2

# cat_df$Year <- as.numeric(cat_df$Year)
# cat_df$Population <- as.numeric(cat_df$Population)
# cat_df$OptTemp_Obs <- as.numeric(cat_df$OptTemp_Obs)
```

```{r}
skimr::skim(cat_df)
round(prop.table(table(cat_df$c_leish_incidence)), 2)
```

```{r}
str(cat_df)
```

Partitioning Data into Training and Test Sets
```{r}
set.seed(1)
library(rsample)
data_split <- initial_split(cat_df, strata = "c_leish_incidence", prop = 0.8)

training <- training(data_split) %>%
  group_by(c_leish_incidence) %>%
  sample_n(size = 3044)
testing <- testing(data_split) 

x_train <- as.matrix(training[,-2])
x_test <- as.matrix(testing[,-2])

y_train <- as.vector(as.numeric(training$c_leish_incidence) - 1)
y_test <- as.vector(as.numeric(testing$c_leish_incidence) - 1)
 
# training <- subset(training, select = -Cutaneous.Leishmaniasis)
# testing <- subset(testing, select = -Cutaneous.Leishmaniasis)

```

```{r}
summary(cat_df)
```

Factorize our target variable
```{r}
training[["c_leish_incidence"]] = factor(training[["c_leish_incidence"]])
```

traincontrol() method:
```{r}
library(caret)
trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
svm_Linear <- train(c_leish_incidence ~., data = training, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneLength = 10)
svm_Radial <- train(c_leish_incidence ~., data = training, method = "svmRadial",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneLength = 10)
```

```{r}
svm_Linear
svm_Radial
```

```{r}
test_pred_Linear <- predict(svm_Linear, newdata = testing)
test_pred_Linear
test_pred_Radial <- predict(svm_Radial, newdata = testing)
test_pred_Radial
```

Predicting Accuracy with Confusion Matrix
```{r}
confusionMatrix(table(test_pred_Radial, testing$c_leish_incidence))
```

We are going to put some values of C using expand.grid() into “grid” dataframe. Next step is to use this dataframe for testing our classifier at specific C values. It needs to be put in train() method with tuneGrid parameter.
```{r}
grid <- expand.grid(C = c(0, 2, 5, 10, 15, 20), sigma = c(0,0.05, 0.1, 0.25,0.5))

svm_Linear_Grid <- train(c_leish_incidence ~., data = training, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneGrid = grid,
tuneLength = 10)
svm_Linear_Grid
plot(svm_Linear_Grid)

svm_Radial_Grid <- train(c_leish_incidence ~., data = training, method = "svmRadial",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneGrid = grid,
tuneLength = 10)
svm_Radial_Grid
plot(svm_Radial_Grid)
```

Let’s try to make predictions using this model for our test set.
```{r}
test_pred_grid <- predict(svm_Linear_Grid, newdata = testing)
test_pred_grid
test_pred_Radial_grid <- predict(svm_Radial_Grid, newdata = testing)
test_pred_Radial_grid
```

Confusion Matrix:
```{r}
confusionMatrix(table(test_pred_Radial_grid, testing$c_leish_incidence))
```










