---
title: "Later Data Multinomial Logistic Regression"
author: "Lyndsey Umsted"
date: '2022-06-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install Package
```{r}
# install.packages("packagename")
```

Libraries
```{r}
require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)
library(dplyr)
```

Importing File into R:
```{r}
aad <- read.csv("Annual_Amazon_Data.csv")
#view(aad)

cutaneous <- aad$Cutaneous.Leishmaniasis
mucosal <- aad$Mucosal.Leishmaniasis
visceral <- aad$Visceral.Leishmaniasis
new_data <- subset(aad, !is.na(cutaneous))
#View(new_data)
#names(new_data)
```


Extracting Later Data and Tidying it:
```{r}
## splitting the data into a before 2014 set and an after 2014 set
library(dplyr)
early_data <- new_data %>%
  filter(Year < 2014)%>%
  select(-c(29:69))
later_data <- new_data %>%
  filter(Year > 2013) %>%
  select(-c(29:69))

#names(early_data)

## removing unnecessary variables

early_data <- early_data %>%
  select(-c("AvgRad"))
later_data <- later_data %>%
  select(-c("StableLights"))

# later_data$OptTemp_Obs <- as.numeric(later_data$OptTemp_Obs)
# later_data$Year <- as.numeric(later_data$Year)
# later_data$Population <- as.numeric(later_data$Population)

later_data_small <- later_data %>%
  select(c("Year", "Population", "Cutaneous.Leishmaniasis", "Dengue_Alb_OptTemp", "Dengue_Aeg_OptTemp", "Chik_Alb_OptTemp", "Chik_Aeg_OptTemp", "Zika_OptTemp", "Malaria_OptTemp", "LST_Day", "LST_Night", "OptTemp_Obs", "NDVI", "EVI", "Precip", "AvgRad", "SWOccurrence")) %>%
  na.omit(later_data_small) 
later_data_t <- later_data_small %>%
  mutate(Cutaneous.Leishmaniasis = (Cutaneous.Leishmaniasis))
```

```{r}
library(dplyr)
summary(later_data_small$Cutaneous.Leishmaniasis[later_data_small$Cutaneous.Leishmaniasis > 0])
cat_df <- subset(later_data_small, later_data_small$Cutaneous.Leishmaniasis > 0)
cat_df$Cutaneous.Leishmaniasis <- cut(later_data_small$Cutaneous.Leishmaniasis[later_data_small$Cutaneous.Leishmaniasis > 0], breaks = c(0, 0.090114445 , 0.876904527, 10^3), labels = c("low", "moderate", "high")) # 25% 75% 100%

# cat_df$label <- NA
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "low"] <- 0
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "moderate"] <- 1
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "high"] <- 2

cat_df$Year <- as.numeric(cat_df$Year)
cat_df$Population <- as.numeric(cat_df$Population)
cat_df$OptTemp_Obs <- as.numeric(cat_df$OptTemp_Obs)
```

```{r}
skimr::skim(cat_df)
round(prop.table(table(cat_df$Cutaneous.Leishmaniasis)), 2)
```

Multinomial Logistic Regression:
```{r}
summary(cat_df)
str(cat_df)
```

Descriptive Analysis
```{r}
library(jmv)

# Use the descritptives function to get the descritptive data
descriptives(cat_df, vars = vars(Year, Population, Cutaneous.Leishmaniasis, LST_Day, LST_Night, OptTemp_Obs, NDVI, EVI, Precip, AvgRad, SWOccurrence), freq = TRUE)
```

```{r}
# To see the crosstable, we need CrossTable function from gmodels package
library(gmodels)
# Build a crosstable between precip and risk level
# CrossTable(cat_df$Precip, cat_df$Cutaneous.Leishmaniasis)
```







Splitting Data into Train and Test Sets:
```{r}
library(caret)

data_split <- createDataPartition(cat_df$Cutaneous.Leishmaniasis, p = 0.75, list = FALSE)
training <- cat_df[data_split,] %>%
  group_by(Cutaneous.Leishmaniasis) %>%
  sample_n(size = 2253)
testing <- cat_df[-data_split,]
```

Setting the reference level
```{r}
# making the reference level "none"
training$Cutaneous.Leishmaniasis <- relevel(training$Cutaneous.Leishmaniasis, ref = "low")
```

Training the Multinomial Classification Model
* To train the model, we will be using multinom function from nnet package. Once the model is trained, then we will use the summary() function to check the model coefficients.

```{r}
require(nnet)
# training the multinomail model
multinom_model <- multinom(Cutaneous.Leishmaniasis ~ ., data = cat_df)

# checking the model
summary(multinom_model)
```

Just like binary logistic regression, we need to convert the coefficients to odds by taking the exponential of the coefficients.
```{r}
exp(coef(multinom_model))
```

The predicted values are saved as fitted.values in the model object. Let’s see the top 6 observations.
```{r}
head(round(fitted(multinom_model), 2))
```

```{r}
# Check the predicted probability for each program
head(multinom_model$fitted.values,20)
head(predict(multinom_model),30)
```

Test the goodness of fit
```{r}
chisq.test(cat_df$Cutaneous.Leishmaniasis,predict(multinom_model))
```

Calculate the Pseudo R-squared
```{r}
# Please takeout the "#" Sign to run the code
# Load the DescTools package for calculate the R square
library(DescTools)
# Calculate the R Square
PseudoR2(multinom_model, which = c("CoxSnell","Nagelkerke","McFadden"))
```
Build a Classification Table
```{r}
# Load the summarytools package to use the classification function
library(summarytools)
# Build a classification table by using the ctable function
ctable <- table(cat_df$Cutaneous.Leishmaniasis,predict(multinom_model))
ctable
```




Predicting and Validating the Model
* To validate the model, we will be looking at the accuracy of the model. This accuracy can be calculated from the classification table.
```{r}
# predicting the values for the train dataset
training$cut.pred <- predict(multinom_model, newdata = training, "class")

# building classifaction table
tab <- table(training$Cutaneous.Leishmaniasis, training$cut.pred)

# calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab))*100,2)
```
Our model accuracy is 76.44% on the training dataset

Predicting the risk level on the Test Dataset
```{r}
# predicting the risk level for the test dataset
testing$cut.pred <- predict(multinom_model, newdata = testing, "class")

# building classifciation table
tab <- table(testing$Cutaneous.Leishmaniasis, testing$cut.pred)
tab

# calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab))*100,2)
```
Our model accuracy is 76.65% on the testing dataset.