---
title: "Support Vector Machine Learning on Later Data"
author: "Lyndsey Umsted"
date: '2022-07-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
load(file = "data/gap_inp")
imp <- gap.inp$ximp
```


```{r}
install.packages("caret")
```

```{r}
library(caret)
```



Extracting Later Data and Tidying it:
```{r}
## splitting the data into a before 2014 set and an after 2014 set
library(dplyr)
early_data <- imp %>%
  filter(Year < 2014)%>%
  select(-c(28:67))
later_data <- imp %>%
  filter(Year > 2013) %>%
  select(-c(28:67))

#names(early_data)

## removing unnecessary variables

early_data <- early_data %>%
  select(-c("AvgRad"))
later_data <- later_data %>%
  select(-c("StableLights"))

# later_data$OptTemp_Obs <- as.numeric(later_data$OptTemp_Obs)
# later_data$Year <- as.numeric(later_data$Year)
# later_data$Population <- as.numeric(later_data$Population)

later_data_small <- later_data %>%
  select(c("Population", "Cutaneous.Leishmaniasis", "Dengue_Alb_OptTemp", "Dengue_Aeg_OptTemp", "Chik_Alb_OptTemp", "Chik_Aeg_OptTemp", "Zika_OptTemp", "Malaria_OptTemp", "LST_Day", "LST_Night", "OptTemp_Obs", "NDVI", "EVI", "Precip", "AvgRad", "SWOccurrence")) 
later_data_t <- later_data_small %>%
  mutate(Cutaneous.Leishmaniasis = (Cutaneous.Leishmaniasis))
```

```{r}
library(dplyr)
summary(later_data_small$Cutaneous.Leishmaniasis[later_data_small$Cutaneous.Leishmaniasis > 0])
cat_df <- subset(later_data_small, later_data_small$Cutaneous.Leishmaniasis > 0)
cat_df$Cutaneous.Leishmaniasis <- cut(later_data_small$Cutaneous.Leishmaniasis[later_data_small$Cutaneous.Leishmaniasis > 0], breaks = c(0, 0.1364497 , 0.4960133, 10^3), labels = c("low", "moderate", "high")) # 33%, 67%, 100%

# cat_df$label <- NA
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "low"] <- 0
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "moderate"] <- 1
# cat_df$label[cat_df$Cutaneous.Leishmaniasis == "high"] <- 2

# cat_df$Year <- as.numeric(cat_df$Year)
# cat_df$Population <- as.numeric(cat_df$Population)
# cat_df$OptTemp_Obs <- as.numeric(cat_df$OptTemp_Obs)
```

```{r}
skimr::skim(cat_df)
round(prop.table(table(cat_df$Cutaneous.Leishmaniasis)), 2)
```

```{r}
str(cat_df)
```

Partitioning Data into Training and Test Sets
```{r}
set.seed(1)
library(rsample)
data_split <- initial_split(cat_df, strata = "Cutaneous.Leishmaniasis", prop = 0.8)

training <- training(data_split) %>%
  group_by(Cutaneous.Leishmaniasis) %>%
  sample_n(size = 5864)
testing <- testing(data_split) 

x_train <- as.matrix(training[,-2])
x_test <- as.matrix(testing[,-2])

y_train <- as.vector(as.numeric(training$Cutaneous.Leishmaniasis) - 1)
y_test <- as.vector(as.numeric(testing$Cutaneous.Leishmaniasis) - 1)
 
# training <- subset(training, select = -Cutaneous.Leishmaniasis)
# testing <- subset(testing, select = -Cutaneous.Leishmaniasis)

```

```{r}
summary(cat_df)
```

Factorize our target variable
```{r}
training[["Cutaneous.Leishmaniasis"]] = factor(training[["Cutaneous.Leishmaniasis"]])
```

traincontrol() method:
```{r}
trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
svm_Linear <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneLength = 10)
svm_Radial <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmRadial",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneLength = 10)
```

```{r}
svm_Linear
svm_Radial
```

```{r}
test_pred_Linear <- predict(svm_Linear, newdata = testing)
test_pred_Linear
test_pred_Radial <- predict(svm_Radial, newdata = testing)
test_pred_Radial
```

Predicting Accuracy with Confusion Matrix
```{r}
confusionMatrix(table(test_pred_Radial, testing$Cutaneous.Leishmaniasis))
```

We are going to put some values of C using expand.grid() into “grid” dataframe. Next step is to use this dataframe for testing our classifier at specific C values. It needs to be put in train() method with tuneGrid parameter.
```{r}
grid <- expand.grid(C = c(0, 2, 5, 10, 15, 20), sigma = c(0,0.05, 0.1, 0.25,0.5))

svm_Linear_Grid <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmLinear",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneGrid = grid,
tuneLength = 10)
svm_Linear_Grid
plot(svm_Linear_Grid)

svm_Radial_Grid <- train(Cutaneous.Leishmaniasis ~., data = training, method = "svmRadial",
trControl=trctrl,
preProcess = c("center", "scale"),
tuneGrid = grid,
tuneLength = 10)
svm_Radial_Grid
plot(svm_Radial_Grid)
```

Let’s try to make predictions using this model for our test set.
```{r}
test_pred_grid <- predict(svm_Linear_Grid, newdata = testing)
test_pred_grid
test_pred_Radial_grid <- predict(svm_Radial_Grid, newdata = testing)
test_pred_Radial_grid
```

Confusion Matrix:
```{r}
confusionMatrix(table(test_pred_Radial_grid, testing$Cutaneous.Leishmaniasis))
```










